---
title: "Ensemble Learning in R for Continuous Response"
author: "Srijit Mukherjee"
date: "27 July 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

* library(ISLR)
* library(MASS)
* library(tree)
* library(randomForest)
* library(gbm)
* library(adabag)
* library(caret)
* library(corrplot)

```{r,  include=FALSE}
library(ISLR)
library(MASS)
library(tree)
library(randomForest)
library(gbm)
library(adabag)
library(caret)
library(corrplot)
```

We will be using the data set Boston Data Set in ISLR package. Let's explore the data.

# Dataset: Boston

This dataset contains information collected by the U.S Census Service concerning housing in the area of Boston Mass.

There are 14 attributes in each case of the dataset. They are:


* CRIM - per capita crime rate by town
* ZN - proportion of residential land zoned for lots over 25,000 sq.ft.
* INDUS - proportion of non-retail business acres per town.
* CHAS - Charles River dummy variable (1 if tract bounds river; 0 otherwise)
* NOX - nitric oxides concentration (parts per 10 million)
* RM - average number of rooms per dwelling
* AGE - proportion of owner-occupied units built prior to 1940
* DIS - weighted distances to five Boston employment centres
* RAD - index of accessibility to radial highways
* TAX - full-value property-tax rate per $10,000
* PTRATIO - pupil-teacher ratio by town
* B - 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town
* LSTAT - % lower status of the population
* MEDV - Median value of owner-occupied homes in $1000's

```{r}
head(Boston)
str(Boston)
```


# Goal

Predict Price (MEDV)

## Train and Test Data Set

```{r}
train = sample (1:nrow(Boston), nrow(Boston)/2)
boston.test=Boston[-train ,"medv"]
```


We will apply the following methods on the same data set and compare them.

* Decision Tree
* Bagging
* Random Forest (Bagging + Decorrelate)
* Boosting

# Decision Tree

Fit the model

```{r}
set.seed(1)
tree.boston = tree(medv ~., Boston ,subset =train)
```

The summary of the Regression Tree

```{r}
summary (tree.boston )
```

The Plot of the Decision Tree

```{r}
plot(tree.boston)
text(tree.boston, pretty = 0)
```

The Plot of the Cross Validation Deviance

```{r}
cv.boston =cv.tree(tree.boston)
plot(cv.boston$size ,cv.boston$dev ,type= "b")
```

After Pruning and setting the number of leaf nodes to be 5.

```{r}
prune.boston = prune.tree(tree.boston, best = "5")
plot(prune.boston)
text(prune.boston ,pretty =0)
```

Performance

```{r}
yhat=predict (tree.boston ,newdata =Boston [-train ,])
plot(yhat ,boston.test)
abline (0,1)
mean((yhat -boston.test)^2)
```


# Bagging

Prediction with Bagging

```{r}
set.seed(2)
bag.boston= randomForest(medv~., data = Boston, subset=train, mtry=13,importance =TRUE)
yhat.bag = predict(bag.boston , newdata = Boston[-train ,])
mean((yhat.bag - boston.test)^2)
```

Visualize the performance of the Bagging

```{r}
summary(bag.boston)
plot(yhat.bag, boston.test)
abline (0,1)
```

Change in the number of trees: Performance Measure

```{r}
bag.boston= randomForest(medv~., data = Boston, subset=train, mtry=13,importance =TRUE, ntree = 25)
yhat.bag = predict (bag.boston , newdata = Boston[-train ,])
mean((yhat.bag - boston.test)^2)
```


# Random Forest

```{r}
corrplot(cor(Boston))
```

Change in the random feature selection

```{r}
rf.boston= randomForest(medv~., data = Boston, subset=train, mtry=6,importance =TRUE, ntree = 25)
yhat.bag = predict (rf.boston , newdata = Boston[-train ,])
mean((yhat.bag - boston.test)^2)
```

Variable Importance Plot

```{r}
importance(rf.boston)
varImpPlot (rf.boston)
```

How the number of trees result in the error?

```{r}
error_ntree = NULL

for (n in 1:100)
{
rf.boston= randomForest(medv~., data = Boston, subset=train, mtry=6, importance =TRUE, ntree = n)
yhat.bag = predict (rf.boston , newdata = Boston[-train ,])
error_ntree = c(error_ntree, mean((yhat.bag - boston.test)^2))
}
plot(error_ntree)
```


How the number of predictor selection result in the error?

```{r}
error_mtry = NULL

for (n in 1:(ncol(Boston)-1))
  
{
rf.boston= randomForest(medv~., data = Boston, subset=train, mtry=6, importance =TRUE, ntree = n)
yhat.bag = predict (rf.boston , newdata = Boston[-train ,])
error_mtry = c(error_mtry, mean((yhat.bag - boston.test)^2))
}
plot(error_mtry)
```



# Gradient Boosting

```{r}
set.seed (3)
boost.boston =gbm(medv ~ ., data= Boston[train ,], distribution = "gaussian", n.trees =5000 , interaction.depth =4)
summary (boost.boston )
par(mfrow =c(1,2))
plot(boost.boston ,i="rm")
plot(boost.boston ,i="lstat")
yhat.boost=predict (boost.boston ,newdata =Boston [-train ,],n.trees =5000)
mean(( yhat.boost -boston.test)^2)
```
